Reproducible Research in R: A tutorial on how to do the same thing more than once

### 2021-09-27 manuscript.Rmd

Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
Preregistrations, however, have several shortcomings.
One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

The workflow described in this paper facilitates a rigorous solution to this problem:
Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
We term this approach of writing and publishing code at the preregistration stage *preregistration as code (PAC)*.
PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
Moreover, it removes the necessity to write two separate documents:
A preregistration *and* a manuscript.
For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
This already includes most typical sections, such as Introduction, Methods, and Results.
These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
Once the preregistration is submitted and real data have been collected or made available,
the document can be reproduced with a single command, thus updating the Results to their final version.

@@

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, the PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
By contrast, filling out a preregistration template is a novel type of scientific writing for many.
Second, the PAC is a tool for study planning.
When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
A PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
Decisions that do not lend themselves to formulation in code, e.g. visual inspection, must still be described verbally or be treated as noted in the section [Deviating from the preregistration and exploration].
With PAC researchers can precisely calibrate how much they want to explore.
When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.

## Deviating from the preregistration and exploration

We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
To make the comparison even more accessible, we highly recommend that the authors review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.

Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
In fact, we expect that most PACs will require changes after data have been collected.
Often deviations provide an opportunity to learn from the unexpected.
Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].

Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
However, after fitting the model on the actual data, both constructs exhibit high collinearity.
The authors decide that for their purpose, these variables are interchangeable and remove one.
For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
Another researcher, however, who, e.g. hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
With transparent reporting, it is left to the reader to make up their own judgement.

[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.

Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning. <!--citation? reproducibility in machine learning research? maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models? -->
These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
We specifically recommend PAC for those exploratory statistical models.
The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.

@@

This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects including reproducible manuscripts.
These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
We first demonstrated how to create a reproducible research project.
Then, we illustrated how such a project can be reproduced - either by the original author and/or collaborators, or by a third party.
We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
such as the preregistration as code (PAC).
A PAC incorporates all planned analysis code into a reproducible manuscript that already renders the expected results.
These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
The entire manuscript, including planned analyses, is preregistered.
This way, all researcher's degrees of freedom are eliminated.
Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.

The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
This analysis plan may contain data-dependent decisions to increase flexibility of research plans.

>>>>

Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
Preregistrations, however, have several shortcomings.
One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

The workflow described in this paper facilitates a rigorous solution to this problem:
Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
We term this approach of writing and publishing code at the preregistration stage *preregistration as code (PAC)*.
PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
Moreover, it removes the necessity to write two separate documents:
A preregistration *and* a manuscript.
For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
This already includes most typical sections, such as Introduction, Methods, and Results.
These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
Once the preregistration is submitted and real data have been collected or made available,
the document can be reproduced with a single command, thus updating the Results to their final version.
Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time.
We can only recommend PAC when the dangers to reproducibility we described above are eliminated and a state of automated reproducibility is reached.

@@

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, the PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
By contrast, filling out a preregistration template is a novel type of scientific writing for many.
Second, the PAC is a tool for study planning.
When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
A PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
With PAC researchers can precisely calibrate how much they want to explore.
When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
and reference it in the manuscript.

Note that preregistration, or PAC, does not preclude exploratory analyses.
Such analyses should merely be explicitly declared as such in the manuscript.

@@

This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects including reproducible manuscripts.
These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
We first demonstrated how to create a reproducible research project.
Then, we illustrated how such a project can be reproduced - either by the original author and/or collaborators, or by a third party.
We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
such as the preregistration as code (PAC).
A PAC incorporates all planned analysis code into a reproducible manuscript that already renders the expected results.
These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
The entire manuscript, including planned analyses, is preregistered.
This way, all researcher's degrees of freedom are eliminated.
Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
Because the same dynamic document serves as preregistration and final manuscript, researchers, reviewers, and editors only have to revise one document.
To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
This analysis plan may contain data-dependent decisions to increase flexibility of research plans.
